\pkg{partykit}: A Modular Toolkit for Recursive Partytioning in \proglang{R}

### 2014-09-05 pkg/partykit/inst/mloss/mloss2.Rnw
rewritten discussion
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@603 255f4249-2263-4417-b1ea-8498779e27a7

%% header declarations
\author{Torsten Hothorn \\ \addr{Universit\"at Z\"urich} \\ \addr{E-mail: Torsten.Hothorn@R-project.org} \AND
        Achim Zeileis \\ \addr{Universit\"at Innsbruck} \\ \addr{E-mail: Achim.Zeileis@R-project.org}}
\title{\pkg{partykit}: A Modular Toolkit for Recursive Partytioning in \proglang{R}}
\editor{(submission for MLOSS track, related to JMLR-07-240-1 that was handled by Cheng Soon Ong)}

@@

\begin{table}
\centering
\begin{tabular}{llll}
\hline
Algorithm  & Software                                                & Object class       & Original reference \\ \hline
CART/RPart & \code{rpart::rpart} + \code{as.party}                   & \code{constparty}  & \cite{Breiman+Friedman+Olshen:1984} \\
C4.5/J4.8  & \pkg{Weka}/\code{RWeka::J48} + \code{as.party}          & \code{constparty}  & \cite{Quinlan:1993} \\
QUEST      & \proglang{SPSS}/\pkg{AnswerTree} + \code{pmmlTreeModel} & \code{simpleparty} & \cite{Loh+Shih:1997} \\
CTree      & \code{ctree}                                            & \code{constparty}  & \cite{Hothorn+Hornik+Zeileis:2006} \\
MOB        & \code{mob}, \code{lmtree}, \code{glmtree}, \dots        & \code{modelparty}  & \cite{Zeileis+Hothorn+Hornik:2008} \\
EvTree     & \code{evtree::evtree}                                   & \code{constparty}  & \cite{evtree} \\ \hline
\end{tabular}
\caption{\label{tab:overview} Selected tree algorithms than can be interfaced through \pkg{partykit}. The
  software column lists external software, \proglang{R} functions from other packages (with \code{::} syntax)
  and from \pkg{partykit}.}
\end{table}

While originally an important goal for the development of \pkg{partykit} was to
provide infrastructure for the authors' own tree induction algorithms
CTree and MOB,
the design was very careful to separate as much functionality as possible into
more general classes that are useful for a far broader class of trees.
In particular, to be able to print/plot/predict different trees in a unified way,
there are so-called coercion functions for transforming trees learned in other
software packages (inside and outside of \proglang{R}) to the classes provided
by \pkg{partykit}. Specifically, tree objects learned by \fct{rpart}
(\citealp{rpart}, implementing CART, \citealp{Breiman+Friedman+Olshen:1984})
and by \fct{J48} from \pkg{RWeka} (\citealp{RWeka}, interfacing \pkg{Weka}'s J4.8
algorithm for C4.5, \citealp{Quinlan:1993}) can be coerced by \fct{as.party}
to the same object class \class{constparty}. This is a general class that can
in principle represent all the major classical tree algorithms with constant
fits in the terminal nodes. Also, the same class is employed for conditional
inference trees (CTree) that can be learned with the \fct{ctree} function
directly within \pkg{partykit} or evolutionary trees from package \pkg{evtree}
\citep{evtree}.

@@

Package \pkg{partykit} provides a toolkit for trees in \proglang{R} that gives
emphasis to flexibility and extensibility. The infrastructure is easily
accessible and written in plain \proglang{R}. Nevertheless speed is not a critical
concern because (a)~lean data structures are employed and (b)~the time-intensive
part in tree computations is typically inferring the tree rather than
representing/printing/plotting it. Hence, the two functions for inferring trees
either have computationally intensive parts in \proglang{C} (\fct{ctree})
or build on \proglang{R}'s fitting functions (\fct{mob}). For benchmarks of
their predictive performance, see their original references.
Regarding memory requirements, the \class{party} trees are very lean but can
optionally be used to store any information. The recursive node structure
itself is extremely lightweight and optionally zero rows from the learning data
are sufficient, which future versions will exploit for storing forests.

>>>>

%% header declarations
\author{Torsten Hothorn \\ \addr{Universit\"at Z\"urich} \\ \addr{E-mail: Torsten.Hothorn@R-project.org} \AND
        Achim Zeileis \\ \addr{Universit\"at Innsbruck} \\ \addr{E-mail: Achim.Zeileis@R-project.org}}
\title{\pkg{partykit}: A Modular Toolkit for Recursive Partytioning in \proglang{R}}
\editor{Cheng Soon Ong (MLOSS track, JMLR-14-133)}

@@

\begin{table}
\centering
\begin{tabular}{llll}
\hline
Algorithm  & Software implementation                                 & Object class       & Original reference \\ \hline
CART/RPart & \code{rpart::rpart} + \code{as.party}                   & \code{constparty}  & \cite{Breiman+Friedman+Olshen:1984} \\
C4.5/J4.8  & \pkg{Weka}/\code{RWeka::J48} + \code{as.party}          & \code{constparty}  & \cite{Quinlan:1993} \\
QUEST      & \proglang{SPSS}/\pkg{AnswerTree} + \code{pmmlTreeModel} & \code{simpleparty} & \cite{Loh+Shih:1997} \\
CTree      & \code{ctree}                                            & \code{constparty}  & \cite{Hothorn+Hornik+Zeileis:2006} \\
MOB        & \code{mob}, \code{lmtree}, \code{glmtree}, \dots        & \code{modelparty}  & \cite{Zeileis+Hothorn+Hornik:2008} \\
EvTree     & \code{evtree::evtree}                                   & \code{constparty}  & \cite{evtree} \\ \hline
\end{tabular}
\caption{\label{tab:overview} Selected implementations of tree algorithms that can be interfaced through \pkg{partykit}. The
  second column lists external software, \proglang{R} functions from other packages (with \code{::} syntax)
  and from \pkg{partykit}.}
\end{table}

While originally an important goal for the development of \pkg{partykit} was to
provide infrastructure for the authors' own tree induction algorithms
CTree and MOB,
the design was very careful to separate as much functionality as possible into
more general classes that are useful for a far broader class of trees.
In particular, to be able to print/plot/predict different trees in a unified way,
there are so-called coercion functions for transforming trees learned in other
software packages (inside and outside of \proglang{R}) to the classes provided
by \pkg{partykit}. Specifically, tree objects learned by \fct{rpart}
(\citealp{rpart}, implementing CART, \citealp{Breiman+Friedman+Olshen:1984})
and by \fct{J48} from \pkg{RWeka} (\citealp{RWeka}, interfacing \pkg{Weka}'s J4.8
algorithm for C4.5, \citealp{Quinlan:1993}) can be coerced by \fct{as.party}
to the same object class \class{constparty}. This is a general class that can
in principle represent all the major classical tree types with constant
fits in the terminal nodes. Also, the same class is employed for conditional
inference trees (CTree) that can be learned with the \fct{ctree} function
directly within \pkg{partykit} or evolutionary trees from package \pkg{evtree}
\citep{evtree}.

@@

Package \pkg{partykit} provides a toolkit for trees in \proglang{R} that gives
emphasis to flexibility and extensibility. The infrastructure is easily
accessible and accompanied by detailed manual pages and package vignettes.
The package facilitates the implementation of new algorithms or interfacing
other software by providing common building blocks for computing on trees
(representation, printing, plotting, predictions, etc.). Using these building
blocks developers of tree software can focus on implementing the learning
algorithm (selection of variables and split points, stopping criteria, pruning, etc.).
The package also provides functions for inferring trees where the computationally
intensive parts are either in \proglang{C} (\fct{ctree}) or employ \proglang{R}'s
fitting functions (\fct{mob}). The simple and lean base classes that separate
data and tree structure are also appealing for storing forests which are planned
for future versions of the package. Users and developers that have questions or
comments about the package can either contact the maintainers or use the forum
on \proglang{R}-Forge at \url{https://R-Forge.R-project.org/forum/forum.php?forum_id=852}.
